### 2009-10-21 latex/paper1/contents.tex
versao submetida para o csmr 2010

Large software systems need to be decomposed in modules in order to be
effectively maintained by groups of developers working concurrently. Software
clustering algorithms
\cite{Mancoridis1998,Anquetil1999,Tzerpos2000,Andritsos2005}, also known as
architecture recovery algorithms, can automatically find a suitable
decomposition by analyzing the dependencies between components in a software
system (e.g., classes in object-oriented systems) and then grouping together
highly interdependent components.

Although many software clustering algorithms have been proposed, there is little
empirical evaluation regarding the external quality of the decompositions they
found. One important test for a software clustering algorithm is to apply it to
a collection of software systems with known reference decompositions made by
experienced developers \cite{Anquetil1999}. It is expected that a good algorithm
finds decompositions that are similar to reference decompositions, which can be
measured by a similarity metric such as MoJo \cite{Tzerpos1999} or EdgeSim
\cite{Mitchell2001}.

@@

An alternative approach is to run the algorithms on synthetic, i.e., computer-
generated, component dependency networks with built-in reference decompositions.
This approach enables one to test algorithms with arbitrarily large samples in a
controlled manner. Nevertheless, this approach is often overlooked, mainly
because networks synthesized by naïve random network models are very dissimilar
from real component dependency networks.

@@

Network theory has been applied to class dependency networks, represented as
directed unweighted graphs, and it was shown that they are also scale-free
\cite{Myers2003,Valverde2002}. Thus, the study of the structure of
object-oriented software systems can benefit from the vast literature available
about scale-free networks. In the remaining of this paper, we will use the
expression ``software network'' to refer to a network of static dependencies
between classes of a software system, in contrast to networks studied in other
domains.

Scale-free network models are algorithms that were proven, either formally or
empirically, to synthesize networks that are scale-free. In this section we
present three models that synthesize scale-free networks with built-in modular
decompositions: BCR+, CGW, and LF. The networks are represented as directed
unweighted graphs with modules, which makes them suitable for testing software
clustering algorithms.

@@

The BCR+ model is a growth model, meaning that the network is created vertex by
vertex, growing from an initial network. It can, therefore, simulate the
evolution of a software network. Moreover, it can simulate the evolution of a
software system subject to constraints in module interaction, as is the case
with top-down system design methodologies.

@@

@@

\begin{figure*}[!t]
\centering
\includegraphics[width=1.0\textwidth]{triads}
\caption{The 13 network triads.}
\label{fig:triads}
\end{figure*}

@@

% Word adjacency network: word Y follows word X => X->Y
\begin{figure}[!t]
\center
\includegraphics{tcp}
\caption{Triad concentration profiles (TCP) for two networks. On the left,
network extracted from the software system JabRef (see the Appendix). On the
right, word adjacency network for the Japanese language \cite{Milo2004}.}
\label{fig:profiles}
\end{figure}

@@

We have shown empirically that network models found in the literature can
synthesize networks that resemble the network of dependencies between classes in
object-oriented systems. This result supports the use of synthetic networks in
the evaluation of software clustering algorithms.
%that operate on class dependency networks.

>>>>

Large software systems need to be decomposed in modules in order to be
effectively maintained by groups of developers working concurrently. Software
clustering algorithms
\cite{Mancoridis1998,Anquetil1999,Tzerpos2000,Andritsos2005}, also known as
architecture recovery algorithms, find suitable decompositions by analyzing the
dependencies between components in a software system (e.g., classes in
object-oriented systems) and then grouping together highly interdependent
components.

Although many software clustering algorithms have been proposed, there is little
empirical evaluation regarding the external quality of the decompositions they
find. One important test for a software clustering algorithm is to apply it to a
collection of software systems with known reference decompositions made by
experienced developers \cite{Anquetil1999}. It is expected that a good algorithm
finds decompositions that are similar to reference decompositions, which can be
measured by a similarity metric such as MoJo \cite{Tzerpos1999} or EdgeSim
\cite{Mitchell2001}.

@@

An alternative approach is to run the algorithms on synthetic, i.e., computer-
generated, component dependency networks with built-in reference decompositions.
This approach enables one to test algorithms with arbitrarily large samples in a
controlled manner. Nevertheless, this approach is often overlooked, mainly
because networks synthesized by naive random network models are very dissimilar
from real component dependency networks.

@@

Network theory has been applied to class dependency networks, represented as
directed unweighted graphs, and it was shown that they are also scale-free
\cite{Myers2003,Valverde2002}. Thus, the study of the structure of
object-oriented software systems can benefit from the vast literature available
about scale-free networks.

Scale-free network models are algorithms that can be proven, either formally or
empirically, to synthesize networks that are scale-free. In this section we
present three models that synthesize scale-free networks with built-in modular
decompositions: BCR+, CGW, and LF. The networks are represented as directed
unweighted graphs with modules, which makes them suitable for testing software
clustering algorithms.

@@

The BCR+ model is a growth model, meaning that the network is created vertex by
vertex, growing from an initial network. It can, therefore, simulate the
evolution of a software system. Moreover, it can simulate the evolution of a
software system subject to constraints in module interaction, as is the case
with top-down system design methodologies.

@@

In the remaining of this paper, we will use the expression ``software network''
to refer to a network of static dependencies between classes of a software
system, in contrast to networks studied in other domains.

@@

\begin{figure}[t]
\centering
\includegraphics{triads}
%[width=1.0\textwidth]{triads}
\caption{The 13 network triads.}
\label{fig:triads}
\end{figure}

@@

% Word adjacency network: word Y follows word X => X->Y
\begin{figure}[t]
\center
\includegraphics{tcp}
\caption{Triad concentration profiles (TCP) for two networks. On the left,
network extracted from the software system JabRef (see the Appendix). On the
right, word adjacency network for the Japanese language \cite{Milo2004}.}
\label{fig:profiles}
\end{figure}

@@

We have shown empirically that network models found in the literature can
synthesize networks that resemble the network of static dependencies between
classes in object-oriented systems. This result supports the use of synthetic
networks in the evaluation of software clustering algorithms.
%that operate on class dependency networks.
